\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-18 paper/main.tex
.

Both approaches may be used in conjunction with network inference
\cite{Chen2015}. Biological networks used for DTI prediction include
protein--protein interaction networks \cite{Feng2017, Lee2018} and
networks including several other types of biological relations,
including similarity between represented entities
\cite{Gottlieb2011}. Network-based DTI prediction methods use the
guilt-by-association principle \cite{Oliver2000} and assume that a
protein is a likely target for a drug if many of the protein's
neighbors in the interaction network are targets of the drug
\cite{Gillis2012}. Network-based methods have been applied
successfully to DTI prediction. However, if DTIs are taken as direct
physical interactions between a drug and protein, it remains an
unresolved question whether the network-based guilt-by-association
hypothesis is true, or whether an interaction of a drug and protein
dys-regulates several of the protein's interaction partners, and
therefore resulting in effects that are not direct interactions but
only downstream consequences of an interaction.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. As this is quite common for DTI datasets, we do not omit but impute the $MicroAUC$
interpolating linearly for those entities, by using the accuracy for this
subset:

\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Note that drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

Further, we choose AUROC over area under precision recall curve (AUPRC) as the latter is prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases, undermining cross-approach evaluation.\todo[inline]{More Information? Right place?}

>>>>

Both approaches may be used in conjunction with network inference
\cite{Chen2015}. Biological networks used for DTI prediction include
protein--protein interaction networks \cite{Feng2017, Lee2018} and
networks including several other types of biological relations,
including similarity between represented entities \cite{Ding2013,
  Gottlieb2011}. Network-based DTI prediction methods use the
guilt-by-association principle \cite{Oliver2000} and assume that a
protein is a likely target for a drug if many of the protein's
neighbors in the interaction network are targets of the drug
\cite{Gillis2012}. Network-based methods have been applied
successfully to DTI prediction. However, if DTIs are taken as direct
physical interactions between a drug and protein, it remains an
unresolved question whether the network-based guilt-by-association
hypothesis is true, or whether an interaction of a drug and protein
dys-regulates several of the protein's interaction partners, and
therefore resulting in effects that are not direct interactions but
only downstream consequences of an interaction.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. As this is quite common for DTI datasets, we do not
omit but impute the $MicroAUC$ interpolating linearly for those
entities using the accuracy for this subset:
\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Note that drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

Further, we choose the AUROC as primary metric to compare different
methods, over area under precision recall curve (AUPRC); AUPRC is
prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases,
undermining cross-approach evaluation.
